---
title: "data_analysis"
author: "Timotius Marselo"
date: "2024-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(corrplot)
library(lsr)
library(glmnet)
library(scales)
```

```{r}
target <- "experienced_maori"
cat_features <- c("travelled_alone",  "travelled_with_child_under_15", 
                 "travelled_with_other_above_15", "drive_yourself", "qtr", 
                 "country_of_residence_group", "gender", "first_nz_trip",
                 "purpose_of_visit_main", "visited_island")
num_features <- c("no_days_in_nz", "age_range", "treated_spend", "spend_per_day")

df <- read.csv("merged_data.csv") %>% 
  mutate_at(cat_features, as.factor) %>%
  mutate_at(num_features, as.numeric)

head(df)
```

```{r}
# set the levels to improve interpretability of the model later
# for target we set "Yes" as the first level, so it will be the positive class
df$experienced_maori <- factor(df$experienced_maori, levels = c("Yes", "No"))
df$travelled_alone <- factor(df$travelled_alone, levels = c("No", "Yes"))
df$travelled_with_child_under_15 <- factor(df$travelled_with_child_under_15, levels = c("No", "Yes"))
df$travelled_with_other_above_15 <- factor(df$travelled_with_other_above_15, levels = c("No", "Yes"))
df$drive_yourself <- factor(df$drive_yourself, levels = c("No", "Yes"))
df$qtr <- factor(df$qtr, levels = c(1, 2, 3, 4))  # Use ordered = TRUE if it's a ranking
df$country_of_residence_group <- factor(df$country_of_residence_group, 
                                        levels = c("Oceania", "Asia", "Africa and Middle East", "Americas", "Europe"))
df$gender <- factor(df$gender, levels = c("Female", "Male"))  
df$first_nz_trip <- factor(df$first_nz_trip, levels = c("No", "Yes"))
df$purpose_of_visit_main <- factor(df$purpose_of_visit_main, 
                                   levels = c("Business", "Visiting friends / relatives", "Other", "Holiday / vacation", "Education"))  
df$visited_island <- factor(df$visited_island, levels = c("South", "North", "Both"))
```

```{r}
# check for duplicated response id
df %>%
  group_by(response_id) %>%
  filter(n() > 1)
```

```{r}
# check for missing values
df %>%
  is.na() %>%
  colSums()
```

```{r}
# check distribution of numerical columns
for (col in num_features) {
  print(ggplot(df, aes_string(col)) + 
          geom_histogram() +
          theme_minimal() +
          labs(title = paste("Distribution of", col)))
}

```
no_days_in_nz, treated_spend and spend_per_day are right skewed, while age_range is bimodal.

```{r}
# check distribution of categorical columns
for (col in c(cat_features, target)) {
  print(ggplot(df, aes_string(col)) + 
          geom_bar(aes(y = after_stat(prop), group=1)) +
          coord_flip() +
          theme_minimal() +
          labs(title = paste("Distribution of", col)))
}
```
There some low value categories in travelled_with_child_under_15 (Yes), country_of_residence_group (Africa and Middle East), purpose_of_visit_main (Business, Education and Other)

```{r}
# visualize relationship between numerical variables and experienced_maori
# using boxplot and density plot
for (col in num_features) {
  print(ggplot(df, aes_string("experienced_maori", col)) + 
          geom_boxplot() +
          labs(title = paste("Distribution of", col, "by experienced_maori")))
          
}

for (col in num_features) {
  print(ggplot(df, aes_string(col, fill = "experienced_maori")) + 
          geom_density(alpha = 0.5) +
          labs(title = paste("Density plot of", col, "by experienced_maori")))
}
```

There seems to be some difference in the distribution of no_days_in_nz, treated_spend and spend_per_day between experienced and non-experienced Maori.

```{r}
# visualize relationship between categorical variables and experienced_maori
# using stakced bar chart with proportion

for (col in cat_features) {
  print(ggplot(df, aes_string(col, fill = "experienced_maori")) + 
          geom_bar(position = "fill") +
          coord_flip() +
          labs(title = paste("Distribution of", col, "by experienced_maori"))) 
}
```

We can see difference in proportion of experienced and non-experienced Maori in almost all categorical variables. The difference by gender, travelled_with_child_under_15, and qtr seems less pronounced.

```{r}
# check for multicollinearity between num features and
# relationship between num features with target using Pearson correlation
df <- df %>% 
  mutate(ohe_experinced_maori = ifelse(experienced_maori == "Yes", 1, 0))

num_var_cor <- cor(df[, c(num_features, "ohe_experinced_maori")]) 
corrplot(num_var_cor, addCoef.col = "black", number.cex = 0.7, cl.cex = 0.7, tl.cex=0.7,cex.main=0.8, diag=FALSE, type="lower", title="Correlation Between Numerical Features and Target", mar=c(0, 0, 1, 0))

df <- df %>% 
  select(-ohe_experinced_maori)
```
The pairwise numerical features has correlation less than 0.7, so we assume no significant multicollinearity. We will drop age_range and spend_per_day for modelling because they have low correlation with the target variable.
```{r}
num_features <- setdiff(num_features, c("age_range", "spend_per_day"))
num_features
```
```{r}
cramers_v_matrix <- function(data) {
  n <- ncol(data)
  cramers_v <- matrix(NA, n, n)
  colnames(cramers_v) <- colnames(data)
  rownames(cramers_v) <- colnames(data)
  
  for (i in 1:n) {
    for (j in i:n) {
      tbl <- table(data[, i], data[, j])
      cramers_v[i, j] <- cramersV(tbl)
      cramers_v[j, i] <- cramers_v[i, j]  
    }
  }
  return(cramers_v)
}
```

```{r}
# check for multicollinearity of categorical variables
result <- cramers_v_matrix(df[c(cat_features, target)])

# plot a correlation matrix
corrplot(result, addCoef.col = "black", number.cex = 0.7, cl.cex = 0.7, tl.cex = 0.7, cex.main=0.7, diag=FALSE, type="lower", title="Cramers V Between Categorical Features and Target", mar=c(0, 0, 1, 0))
```
We can notice that travelled_alone and travelled_with_other_above_15 have a high correlation. This makes sense because if someone is not travelling alone, then they are travelling with others. Also the travelled_with_child_under_15 has 0 cramer's v with the target variable (experienced_maori), which means their relationship is weak. We will drop travelled_alone and travelled_with_child_under_15 for modelling. gender also has low correlation with the target variable, so we will drop it as well.

```{r}
# drop travelled_alone and travelled_with_child_under_15
cat_features <- setdiff(cat_features, c("travelled_alone", "travelled_with_child_under_15", "gender"))

# do log transformation because we have right skewed numerical variables
log_num <- c("no_days_in_nz", "treated_spend")

df_processed <- df %>% 
  mutate_at(log_num, log1p)
```

```{r}
# standardize the numerical variables so ridge regression can be trained properly
df_processed <- cbind(df_processed[, c("experienced_maori", cat_features)], scale(df_processed[, num_features]))
```

```{r}
# plot histogram of numerical variables after log transformation
for (col in num_features) {
  print(ggplot(df_processed, aes_string(col)) + 
          geom_histogram())
}
```

```{r}
library(caret)
set.seed(123)
```

```{r}
# create cross validation setup
train_control <- trainControl(method = "cv", 
                              number = 10,
                              classProbs = TRUE,
                              savePredictions = "all",
                              summaryFunction = twoClassSummary)
```

```{r}
# train ridge regression model
ridge_logreg_model <- train(experienced_maori ~ ., 
               data = df_processed, 
               method = "glmnet",            
               family = "binomial",
               trControl = train_control,
               metric = "ROC",
               tuneGrid = expand.grid(alpha = 0, lambda = c(0, 0.001, 0.01, 0.1, 1, 10)))
```

```{r}
# show the results of the ridge regression model across different lambda values
ridge_logreg_model$results %>% 
  # select(alpha, lambda, ROC, Sens, Spec) %>% 
  arrange(desc(ROC))
```

```{r}
coefs <- coef(ridge_logreg_model$finalModel, s = 0)
coefs <- as.numeric(coefs[-1])
names(coefs) <- rownames(coef(ridge_logreg_model$finalModel, s = 0))[-1] 
```

```{r}
# reverse the sign of the coefficients, to make the "no" target as a baseline
coefs_df <- data.frame(coefs = coefs * -1, features = names(coefs))

coefs_df %>%
  ggplot(aes(x = reorder(features, coefs), y = coefs)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Coefficient for Logistic Regression (lambda = 0)", x = "Features", y = "Coefficient") +
  theme_minimal()
```
```{r}
cv_predictions <- ridge_logreg_model$pred

best_cv_predictions <- cv_predictions %>% 
  filter(lambda == 0) 
```

```{r}
conf_matrix <- confusionMatrix(best_cv_predictions$pred, best_cv_predictions$obs)

cm_table <- conf_matrix$table %>% 
  as.data.frame()
```


```{r}
cm_table %>%
  ggplot(aes(x=Reference, y=Prediction, fill = Freq)) +
    geom_tile(color = "black") +
    geom_text(aes(label = Freq),color="black", size=5) +
    scale_fill_gradient(low = "lightblue", high = "blue") +
    labs(
      title = "Confusion Matrix",
      x = "Actual Class",
      y = "Predicted Class"
    ) +
    theme_minimal()
```
```{r}
# plot feature importance of the ridge regression model
importance <- varImp(ridge_logreg_model, scale = TRUE)
plot(importance, main = "Variable Importance for Ridge Logistic Regression")
```
```{r}
library(pROC)

roc_curve <- roc(best_cv_predictions$obs, best_cv_predictions$Yes)

plot(roc_curve, 
     main = "ROC Curve for Logistic Regression (lambda = 0)", 
     col = "red", 
     lwd = 2)

auc_value <- auc(roc_curve)

legend("bottomright", 
       legend = paste("AUC =", round(auc_value, 3)),
       bty = "n")  
```
```{r}
# calculate fitted values (predicted probabilities)
fitted_values <- as.numeric(best_cv_predictions$Yes)

# Calculate residuals
binary_obs <- ifelse(best_cv_predictions$obs == "Yes", 1, 0)
residuals <- binary_obs - fitted_values
```

```{r}
# Create the binned residual plot
library(arm)
binnedplot(fitted_values, 
           residuals, 
           xlab = "Expected Values", 
           ylab = "Average Residual", 
           main = "Binned Residual Plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")
```

```{r}
# train the random forest model
library(randomForest)
library(caret)

mtry = c(3, 4, 5)
ntree_values <- c(125, 250)
node_size <- c(3, 6, 9)

# loop over ntree values to create models
rf_models <- list()

for (ntree in ntree_values) {
    for (nodesize in node_size) {
        rf_model <- train(
          experienced_maori ~ ., 
          data = df_processed, 
          method = "rf", 
          trControl = train_control, 
          tuneGrid = expand.grid(mtry = mtry),
          metric = "ROC",
          ntree = ntree, 
          nodesize = nodesize,
          importance = TRUE 
        )
        
        # store the model in a list
        rf_models[[paste0("ntree_", ntree, "_nodesize_", nodesize)]] <- rf_model
    }
}
```

```{r}
# iterate through the list of models to extract the results
rf_results_list <- list()

for (rf_model_name in names(rf_models)) {
    current_rf_model <- rf_models[[rf_model_name]]
    
    # extract ROC, sensitivity, specificity for the current model
    current_rf_roc <- current_rf_model$results$ROC  
    current_rf_sens <- current_rf_model$results$Sens
    current_rf_sensSD <- current_rf_model$results$SensSD
    current_rf_spec <- current_rf_model$results$Spec  
    current_rf_specSD <- current_rf_model$results$SpecSD
    
    # extract mtry, ntree, nodesize for the current model
    current_rf_mtry <- current_rf_model$results$mtry
    current_rf_ntree <- current_rf_model$dots$ntree
    current_rf_nodesize <- current_rf_model$dots$nodesize 
    
    # Store results in the list without filtering for max
    rf_results_list[[rf_model_name]] <- data.frame(
        mtry = current_rf_mtry,
        ntree = current_rf_ntree,
        nodesize = current_rf_nodesize,
        ROC = current_rf_roc,
        Sens = current_rf_sens,
        Spec = current_rf_spec,
        SensSD = current_rf_sensSD,
        SpecSD = current_rf_specSD
    )
}
```

```{r}
rf_results_df <- do.call(rbind, rf_results_list)
```

```{r}
rf_results_df %>% 
  arrange(desc(ROC))
```

```{r}
# plot feature importance for the best rf model
best_rf_model <- rf_models[["ntree_250_nodesize_3"]]
importance <- varImp(best_rf_model, scale = FALSE)
plot(importance, main = "Variable Importance for Random Forest")
```

